Spherical harmonic analysis of faceted spheroids identifies shaping strategies and standardisation at Qianshangying (North China)

### 2025-08-14 analysis/paper/paper.qmd
working on the two UMAP figures

p_raw_material <- 
ggplot(df_long %>% 
         mutate(category = case_when(
           category == "cores" ~ "Cores",
           category == "spheroids" ~ "Spheroid",
           category == "total" ~ "Total",
           category == "Polys_and_Multi_cores" ~ "Polys and Multi cores"
           )) %>% 
         filter(category != "Total"), 
       aes(x = material, 
           y = prop, 
           fill = category)) +
  geom_col(position = position_dodge(width = 0.6), width = 0.5) +
  labs(
    x = "",
    y = "Percentage (%)",
    fill = "Technological Type"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
    scale_fill_manual(values = c(
    "Spheroid" = "orange", 
    "Polys and Multi cores" = "#66c2a5",
    "Cores" = "lightgray"
  )) +
  theme_minimal(base_size = 8) +
  theme(
    legend.position = c(0.87, 0.95),              
    legend.justification = c("right", "top"),      
    legend.background = element_rect(fill = alpha("white", 0.6), color = NA),  
    legend.box.background = element_rect(color = "black"),
    legend.box.margin = margin(0, 10, 0, 10)
  )

@@

plot_grid(p_raw_material,
          p_dimensions,
          labels = "AUTO")

@@

# Define the position for the text block
x_pos <- 0.875
y_pos_start <- 1.04
v_spacing <- 0.005 # The vertical gap between lines

@@

>>>>

p_raw_material <- 
ggplot(df_long %>% 
         mutate(category = case_when(
           category == "cores" ~ "Cores",
           category == "spheroids" ~ "Spheroid",
           category == "total" ~ "Total",
           category == "Polys_and_Multi_cores" ~ "Polys and Multi cores"
           )) %>% 
         filter(category != "Total"), 
       aes(x = material, 
           y = prop, 
           fill = category)) +
  geom_col(position = position_dodge(width = 0.6), width = 0.5) +
  labs(
    x = "",
    y = "Percentage (%)",
    fill = "Technological Type"
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
    scale_fill_manual(values = c(
    "Spheroid" = "orange", 
    "Polys and Multi cores" = "#66c2a5",
    "Cores" = "lightgray"
  )) +
  theme_minimal(base_size = 8) +
  theme(
    legend.position = c(0.88, 0.95),              
    legend.justification = c("right", "top"),      
    legend.background = element_rect(fill = alpha("white", 0.6), color = NA),  
    legend.box.background = element_rect(color = "black"),
    legend.box.margin = margin(0, 10, 0, 10)
  )

@@

p_basic_characteristics <- 
plot_grid(p_raw_material,
          p_dimensions,
          labels = "AUTO")

ggsave(filename = here("analysis/paper/figures/basic_artefact_characteristics.png"),
       plot = p_basic_characteristics,
       h = 5, 
       w = 7)

knitr::include_graphics(here("analysis/paper/figures/basic_artefact_characteristics.png"))

@@

# Define the position for the text block
x_pos <- 0.875
y_pos_start <- 1.07
v_spacing <- 0.005 # The vertical gap between lines

@@

```{r}
#| label: fig-UMAP-SPHARM-samples-clusters
#| fig-cap: "A. UMAP plot showing distinct clusters of shapes derived from SPHAR. B UMAP plot of SPHARM-derived shape features colored by types."

library(magick)

p_top <- ggdraw() +
  draw_image(here("analysis/paper/figures/UMAP_SPHARM_clusters.png"))

p_bottom <- ggdraw() +
  draw_image(here("analysis/paper/figures/UMAP_SPHARM_samples.png"))

# use plot_grid() to combine them.
plot_grid(
  p_top,
  p_bottom,
  ncol = 1,
  labels = "AUTO",
  label_size = 10,
  label_fontface = "bold"
)
```
